---
title: "Data 605 Assignment-9"
author: "Deepak Mongia"
date: "April 1, 2019"
output: html_document
---

###Problem 1 - #11 on page 363: The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the nth day of the year. Finn observes that the differences $X_n = Y_{n+1} - Y_n$ appear to be independent random variables with a common distribution having mean $\mu$ = 0 and variance $\sigma ^2$ = 1/4. If $Y_1$ = 100, estimate the probability that $Y_{365}$ is

###(a)  $\ge$ 100

###Answer:
```{r}
mean <- 0
variance <- 1/4
sd <- sqrt(variance)
n <- 364

## calculating x as : (value - mean)/sqrt(n)
x1 <- (100 - 100)/sqrt(n)
pnorm(x1, mean = mean, sd = sd, lower.tail = FALSE)
```

###(a)  $\ge$ 110

###Answer:
```{r}

## calculating x as : (value - mean)/sqrt(n)
x2 <- (110 - 100)/sqrt(n)
pnorm(x2, mean = mean, sd = sd, lower.tail = FALSE)
```

###(a)  $\ge$ 120

###Answer:
```{r}

## calculating x as : (value - mean)/sqrt(n)
x3 <- (120 - 100)/sqrt(n)
pnorm(x3, mean = mean, sd = sd, lower.tail = FALSE)
```

###2. Calculate the expected value and variance of the binomial distribution using the moment generating function.

Answer: PMF for binomial distribution is defined as:
$\binom{n}{x} p^{x} q^{n-x}$

MGF would be:

M(t) = $$\sum_{x=1}^{n} e^{tx} \binom{n}{x} p^{x} q^{n-x}$$
     = $$\sum_{x=1}^{n} \binom{n}{x} {(pe^t)}^x q^{n-x}$$
     = ${(pe^t + q)}^n$
     
For Expected value, taking the first derivative of MGF and then substitute t by 0.
E(x) = np

Similarly
$E(X^2) = n(n-1)p^2 + np$

Var(X) = $E(X^2) - E(X)^2$
= $np(1-p)$

###3. Calculate the expected value and variance of the exponential distribution using the moment generating function.

Answer:
PDF for exponential function = $\lambda e^{-x\lambda}$

$M(t) = \int_{0}^{\infty} e^{tx} \lambda e^{-x\lambda}$

$= \lambda \int_{0}^{\infty} e^{-x(\lambda - t)}$

$= \lambda / {(\lambda - t)}$

$E(X) = 1/\lambda$

$E(X^2) = 2/{\lambda ^2}$

$Var(X) = E(X^2) - E(X)^2$

$= 2/{\lambda ^2} - 1/{\lambda ^2}$

$= 1/{\lambda ^2}$